My title

### 2024-11-02 paper/paper.qmd
Discussion Topics: Swing states
test
model and graphs

# Filter NYT polls and remove entries with NA or empty value for statr
# Select imporntat variables (change as you please)

#& !is.na(state) & state != ""

preddata <- data %>%
  filter(pollster_id == 1424) %>%
  select(display_name, question_id, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date)

print(preddata)
```

@@

#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?
harrisdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    end_date = mdy(end_date)    #is date mutated correctly here 
  ) %>%

  # Step 4: Filter by date and calculate number of respondents
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(
    num_harris = round((pct / 100) * sample_size, 0)
  )

@@

## MODEL 1, BAYESIAN MODEL WITH PREDICTOR STATE

# Define Model
model1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | state)

priors <- normal(0, 2.5, autoscale = TRUE)

bayesian_model1 <- stan_glmer(
  formula = model1,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)

# Plot checks for model 1
pp_check(bayesian_model1)
summary(bayesian_model1)
plot(bayesian_model1, pars = "(Intercept)", prob = 0.95)
print(modelsummary(bayesian_model1))
```

@@

# Define Model 2
model2 <- cbind(num_harris, sample_size - num_harris) ~ (1 | end_date_num) + (1 | state)

bayesian_model2 <- stan_glmer(
  formula = model2,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
```

@@

# Plot checks for Model 2
pp_check(bayesian_model2)
summary(bayesian_model2)
plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
print(modelsummary(bayesian_model2))
```

@@

spline_model_pred <- stan_glm(
  pct ~ ns(end_date_num, df = 5) + state,
  data = harrisdata,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

@@

```{r}

new_data <- data.frame(
  end_date_num = seq(
    min(harrisdata$end_date_num),
    max(harrisdata$end_date_num),
    length.out = 100),
  state = factor("Georgia", levels = levels(harrisdata$state)))

@@

pred_summary <- new_data |>
  mutate(
    pred_mean = colMeans(posterior_preds),
    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975))

ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date_num, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time with Spline Fit") +
  theme_minimal()
```

@@

This paper aims to analyze polling data related to the 2024 election, with a particular emphasis on the swing states. By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions. Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states. Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election.

@@

## First discussion point {#sec-first-point}

@@

## Second discussion point

@@

## Third discussion point

>>>>

# Select imporntat variables (change as you please)

preddata <- data %>%
  filter(numeric_grade >1) %>%
  select(pollster, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date, population, pollscore, numeric_grade) %>%
  filter(!is.na(pollscore)) %>%
  filter(!is.na(state) & state != "")

print(preddata)

@@

#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?
harrisdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    end_date = mdy(end_date)    #is date mutated correctly here
  ) %>%
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(
    num_harris = round((pct / 100) * sample_size, 0)
  )

@@

@@

# Define Model 1
model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)

bayesian_model <- stan_glmer(
  formula = model,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

@@

# Plot checks for Model
pp_check(bayesian_model)
summary(bayesian_model)
plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
print(modelsummary(bayesian_model))
summary(bayesian_model)
```

@@

spline_model_pred <- stan_glm(
  pct ~ ns(end_date_num, df = 5) + state + pollster,
  data = harrisdata,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

@@

# Convert `pollster` to a factor with defined levels if it's not already
harrisdata$pollster <- factor(harrisdata$pollster, levels = unique(harrisdata$pollster))

# Define the new data grid for prediction, ensuring each factor has the correct levels
new_data <- data.frame(
  end_date_num = seq(
    min(harrisdata$end_date_num),
    max(harrisdata$end_date_num),
    length.out = 100
  ),
   pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
  state = factor("Georgia", levels = levels(harrisdata$state)))

@@

pred_summary <- new_data |>
  mutate(
    pred_mean = colMeans(posterior_preds),
    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
  )

ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date_num, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time with Spline Fit") +
   facet_wrap(~ state, scales = "free_y") +
  theme_minimal()
```

@@

In this paper we aim to analyze polling data related to the 2024 election, with a particular emphasis on the swing states. By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions. Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states. Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election.

@@

## Arizona {#sec-arizona}

@@

## Georgia {#sec-georgia}

@@

## Michigan {#sec-michigan}

## Nevada {#sec-nevada}

## North Carolina {#sec-norcar}

## Michigan {#sec-michigan}

## Pennsylvania {#sec-penn}

## Wisconsin {#sec-Wisconsin}
